### 2024-02-15 outputs/paper/paper.qmd
Final Commit

---
title: "Exploring the Shadows of Terror: The Influence of Terrorism on Public Opinion and Democratic Values"
author: 
  - Sinan Ma
  - Yuean Wang
  - Yang Zhou
thanks: "Code and data are available at: https://github.com/Sinanma/Terrorism_And_Citizens-_Political_Attitudes. Replication on Social Science Reproduction platform available at: "
date: today
date-format: long
abstract: "In an era where terrorism intersects with global politics, understanding its repercussions on public attitudes towards democracy is crucial. This paper replicates and extends the data analysis of effects of terrorism on public political opinion, considers various expressions of terrorism and a variety of research methodologies by analyzing findings from previous research. Our findings indicate that terrorism leads to an increase in outgroup hostility and a shift towards political conservatism, and observes variations in rally-'round-the-flag effects. These reactions are influenced by the nature of the terrorist threat and the sociopolitical context. Through this exploration, this paper contributes to the ongoing discourse on the implications of terrorism for democratic societies."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r setup}
#| echo: false
#| warning: false
#| message: false
#| 
library(tidyverse)
library(readxl) # reading excel files
library(ggplot2)
library(ggpubr) # for gg rearrange
library(dplyr)
```

# Introduction

In the 21st century, terrorism has emerged as a pivotal concern in global politics, influencing international relations and societal norms worldwide. The impact of terrorism necessitates comprehensive analysis, particularly regarding its influence on public attitudes towards democracy, outgroup hostility, political conservatism, and national unity. Despite the prevalence of terrorist incidents and their apparent effects on society, a critical gap remains in understanding how these events shape public opinion across different contexts.

This paper builds upon the foundational work of Amélie Godefroidt in 2022, published in the American Journal of Political Science [@Godefroidt2023], which highlighted the significant impact of terrorism on public attitudes. Our study aims to replicate and extend Godefroidt's analysis by examining the effects of various forms of terrorism on public sentiment, from domestic incidents to global terror campaigns. We use a meta-analytical approach, integrating findings from a broad spectrum of research to provide an overview of terrorism's impact on public opinion.

The paper delves into the complex dynamics of terrorism's influence on public political opinion. It extends the discourse beyond conventional analysis to explore how various expressions of terrorism—from domestic to international incidents—affect public attitudes, with a particular focus on outgroup hostility, political conservatism, and rally-'round-the-flag phenomena. Despite a growing body of research in this field, a significant gap persists in understanding the holistic and nuanced effects of different forms of terrorism across diverse sociopolitical contexts. To address this gap, we systematically analyze and integrate findings from prior research, employing a variety of methodological approaches to offer a comprehensive overview of terrorism's impact on public sentiment.

Our findings reveal that terrorism significantly influences public opinion, catalyzing a shift towards outgroup hostility and political conservatism, with varying effects on national unity. These shifts are contingent upon the nature of the terrorist threat and the specific sociopolitical context, underscoring the importance of contextual factors in public responses to terrorism. This research contributes to the ongoing discourse on terrorism's implications for democratic societies, offering empirical evidence of its subtle yet profound impact on public political attitudes. By elucidating the conditions under which terrorism most significantly affects public sentiment, this study aims to inform both policymakers and scholars, guiding the development of strategies to enhance democratic resilience in the face of terrorism.

The paper is structured as follows: it begins with an overview of the methodological framework guiding our analysis in [Data section](#sec-data), followed by a presentation of our findings in [Results section](#sec-data-results). We then discuss these in the context of their broader implications for democracy and conclude with recommendations for future research and policy considerations in [Discussion section](#sec-data).

\newpage

# Data {#sec-data}

## Source
Our study is based on a detailed examination by Amélie Godefroidt published in 2022 in the American Journal of Political Science. This important research looks into how terrorism affects what people think about democracy and politics, using a wide range of examples from small local attacks to large international ones. We aim to do more research based on Godefroidt's work [@Godefroidt2023], using her collection of data from 241 papers, which cover 326 studies and provide 1,733 measurements of how terrorism influences people's political views. This collection of information, gathered through careful research and organization, is the main resource for our study as we explore the complex effects of terrorism in different settings around the world.

## Methodology
Our approach to expanding on the original study involves a detailed process known as meta-analysis. This means we take a closer look at a bunch of studies already done on how terrorism changes people's opinions and combine their results to get a clearer picture. Following the steps outlined by Godefroidt, we gather, sort out, and examine data from various sources, focusing on 1,733 data points that show the relationship between terrorism and public opinion. We make sure all these points can be compared fairly by converting them into a standard format and adjusting them so they all point in the same direction. This makes our analysis strong and trustworthy. We also add new studies to the mix and use more sophisticated statistical methods to handle the data better, aiming to deepen our understanding of terrorism's impact on how people feel about democracy, strangers, and political beliefs in different parts of the world.

We used the following packages: `readxl` [@readxl], `ggplot2` [@ggplot2], `ggpubr` [@ggpubr], `dplyr` [@dplyr] in `R` [@citeR] to create figure and analyze data.

\newpage
# Results {#sec-data}

The results section analyses the research trends and diverse reactions to ideological terrorism, supported by two figures. The analysis begins with a detailed examination of the evolution in terrorism research over the past thirty-five years, highlighted by @fig-terrorism-trend and @fig-measure, which showcases shifts in academic focus, data collection, and publication patterns in response to global terrorism events. Then we analyse on @fig-reactions-to-terrorism, which provides a nuanced view of how terrorism impacts political attitudes and public reactions across different ideologies, study designs, and geographical contexts. Through these visual representations, the section underscores the complexity of terrorism's societal impact and the importance of diversifying research approaches to fully understand its multifaceted nature.

## Comprehensive Review of Research Trends
The set of graphs presented in @fig-terrorism-trend and @fig-measure illustrate a multifaceted view of terrorism studies over the past thirty-five years. Each graph demonstrates different aspects and emphases in the study of terrorism, reflecting shifts in global events, academic focus, and thematic categorization. The results emphasize the reactive nature of terrorism research and the importance of diversifying the types of terrorism studied as well as the measures used to assess its impact.
It is critical to understanding how the field is developing and where may need future attention.

### Temporal Trends in Data Collection and Publication

The Year of Data Collection graph (Panel A in @fig-terrorism-trend) illustrates the frequency of data collection efforts over time. The notable peak in the graph indicates a notable increase in terrorism research following significant global incidents, such as the 9/11 attacks and the rise of the Islamic State (IS). These moments stimulate the academic community to investigate the societal repercussions of terrorism following high-profile incidents. Similarly, the Year of Publications graph (Panel B in @fig-terrorism-trend) demonstrates a concurrent increase in the publication of scholarly works, with notable peaks following major terrorist occurrences. The lag between data collection and publication peaks may indicate the time taken for analysis and peer review. This graph underscores the reactive nature of academic publishing in response to terrorism, the understandings in response to the immediate effects of such incidents and increased potential for saturation of literature after major events.

\newpage
```{r }
#| label: fig-terrorism-trend
#| fig-cap: Summary of Data Included in the Meta-Analysis
#| echo: false
#| warning: false
#| message: false
#| out.width: "90%"  # Adjust to fit the page
#| out.height: "90%"  # Adjust to fit the page

knitr::include_graphics("../../outputs/figures/Figure1.jpeg")
```

### Variability in Terrorism Measures

The Terrorism Measures graph (Panel C in @fig-measure) illustrates the diversity of metrics used to measure the impact terrorism. Acts of violence and threat exposure are the most commonly studied, indicating a research focus on direct experiences and perceptions of terrorism. In the meanwhile, it shows an emerging interest in the affective responses elicited by terrorism, even though fear and anger are expressed less frequently. These findings suggest a broader recognition of terrorism's  effects on economic disruptions, reduce productivity, and damage infrastructure, thereby reflecting the multifaceted societal impacts of terrorist.

### Diversity in Types of Terrorism Studies

The Type of Terrorism graph (Panel D in @fig-measure) illustrates the focus of the academic and policy-making communities. It indicates an over-whelming focus on Islamist terrorism, overshadowing other ideologies and forms of terrorism. This skew suggests a potential blind spot in the literature, where state terror and extreme-right terrorism are relatively under-examined, despite their significance in certain regions and contexts.

```{r }
#| label: fig-measure
#| fig-cap: Summary of Data Included in Terrorism Measures and Terrorism Studies
#| echo: false
#| warning: false
#| message: false
#| out.width: "90%"  # Adjust to fit the page
#| out.height: "90%"  # Adjust to fit the page

knitr::include_graphics("../../outputs/figures/Figure2.jpeg")
```

\newpage
## Diverse Reactions to Ideological Terrorism

We use forest plot in @fig-reactions-to-terrorism to illustrate the relationship between terrorism and political attitudes varies by ideology, study design, sample type, and country context. It demonstrates the relationship between terrorism and political attitudes conditional on various substantive and methodological factors. The plot indicates that Islamist terrorism has a more pronounced effect on outgroup hostility, conservative ideological shifts, and rally effects compared to other ideologies.

We used the following additional packages: `broom` [@broom], `dotwhisker` [@dotwhisker], `cowplot` [@cowplot], `metafor` [@metafor] in `R` [@citeR] to create @fig-reactions-to-terrorism.

### Outgroup Hostility
This graph @fig-reactions-to-terrorism presents the Fisher's Z correlation coefficients for the effect of terrorism on outgroup hostility. It demonstrates that terrorism associated with Islamism tends to have a stronger correlation with outgroup hostility than other ideologies. This could reflect media portrayal and public perception biases, where Islamist are more closely associated with the concept of terrorism. The sample type and country also moderate the correlation, with general population samples showing a more substantial effect than student or convenience samples, and the United States showing a stronger correlation than Israel or other countries. In general, other ideologies, or the no ideological label, provoke less intense reactions, suggesting public sensitivity to the context of terrorist events.

### Conservative Shift
The second panel in @fig-reactions-to-terrorism of the forest plot illustrates the effect of terrorism on conservative ideological shifts. Again, the impact is most significant in the context of Islamist terrorism. The effect size is smaller across studies with other or no specified ideologies. Study design and sample type appear to influence the size of the conservative shift, with cross-sectional designs showing a smaller effect size compared to experimental or other designs. This suggests that the presence of a clear ideological motive in terrorism might be a key in changing public opinion towards conservative policies and viewpoints.

### Rally Effects

The third panel in @fig-reactions-to-terrorism of the forest plot illustrates the impact of terrorism on rally effects, which refers to the increased support for existing leadership following terrorist attacks. The data suggests that rally effects are more pronounced in studies examining Islamist terrorism. When examining the influence of research design, experimental designs showed a more marked rally effect compared to cross-sectional studies. This difference indicates that the controlled environment of an experiment, often simulating a direct terrorism threat, might amplify the unifying sentiment among the participants.

### General Population vs. Targeted Samples

The general population samples showed a more consistent increase in outgroup hostility and conservatism across the board, as opposed to student or convenience samples. This highlights the importance of sample representativeness in measuring accuracy public opinion in response to terrorism.

### Geographic Variations
The country-specific analysis revealed that the United States and Israel, both countries with prolonged experiences of terrorism, showed stronger correlations across all three measures compared to other countries. This may be attributed to the higher frequency of terrorist incidents and the consequent heightened public awareness and reaction in these nations.

### Guilt-by-Association Effects
Lastly, the 'Guilt-by-Association' section of @fig-reactions-to-terrorism provides an intriguing insight into how strong associations with terrorism significantly affect public attitudes, more important than moderate or no associations. This underscores the powerful role of narrative and association in shaping public sentiment post-terrorism.

```{r}
#| label: fig-reactions-to-terrorism
#| fig-cap: Relationship between Terrorism and Political Attitudes Conditional on Substantive and Methodological Factors
#| echo: false
#| warning: false
#| message: false
#| out.width: "90%"  # Adjust to fit the page
#| out.height: "90%"  # Adjust to fit the page
knitr::include_graphics("../../outputs/figures/Figure3.jpeg")
```

\newpage

# Discussion {#sec-data}

Investigating terrorism’s effects through public attitudes towards politics is necessary to detect the social reaction to such thought-provoking disasters. Starting from Amélie Godefroidt’s pioneering study [@Godefroidt2023], this paper takes her insights further and tries to bring in more deterministic factors that may work towards shaping public opinion related to terrorism. However, we all agree on the implications of ethics, limitations, and future hopeful prospects.

## Findings

Our research proves to be only an addition to Godefroidt's theory; however, our findings clearly point to the direct effects of terrorist acts on citizens' political attitudes. One striking feature here is that of hatred for out-group members and political conservatism of the public following such catastrophes. This affirms that terror creates a feeling of insecurity and a need for safety, which is a reason for a more cautious political position. Also, our results put emphasis on the way terrorism is shaped in the influence of terrorism. For instance, ideological violence perpetrated in the name of Islam seems to have a tremendous amplifying impact and induces much stronger responses than violence underpinned by other ideologies. This is in accord with former studies focusing on the specific consequences of Islamist extremism on Western countries that associate the latter with the idea of cultural and national identity. Thus, it is considered a threat to civic and national identity.

## Ethical Implications and Bias

Studying terrorism from an ethical dimension proves it cannot be underestimated. Being such a consequence-causing sensitive part of the topic, and its role in evoking and spreading fear and xenophobia, scientists should handle it with specific techniques. An application of rigorous meta-analysis methodology assures reliable inquiry, lack of harm, and the polysemic exploration of the problem of the correlation between terrorism and political attitudes. Nevertheless, it should be acknowledged that the existence of bias is one of the key challenges that could be connected with the initial studies in our analysis. For example, each victim is likely to be bombarded over the media and end up being associated with Islamist terrorism, which is highly visible in media coverage and public discourse. Albeit, while one dimension of ideological violence may not be accurately portrayed, other forms such as far-right extremism or eco-terrorism may not be sufficiently emphasized. Such dyads are likely to find themselves in an arena in which distorted views will dominate and the situation will be perceived far from its true scale.

## Limitations

In addition to the implemented design in order to realize the highest accuracy of a meta-analysis, we faced several limitations, limiting the range and generality of our findings. Our approach, which heavily depends on secondary data, is unable to control for all the confounding variables that may be influential, such as socio-economic class or psychological characteristics of the individual. Furthermore, the use of published sources may leave room for the ‘file drawer’ effect, wherein studies with positive results are likely to be published while studies with no significant effect are simply left in the drawer.

## Future Research

To rectify these limitations as well as bring clarity into how terrorism changes the political views of a society, there should be an all-around approach to the research. At first, applying primary data collection methods, e.g., surveys or interviews, is a good idea since it can bring you concrete views of how people behave in terrible situations. This will ultimately allow the scientists to enable the research system to explore intricacies and contextual factors which may not be available in the datasets that currently exist. In addition, the issue must be viewed from a wider angle, not being limited only to a single terrorist ideology but covering the broad spectrum of ideologies. Our primary target, however, revolves around Islamist terrorism alone, yet other extreme movements, for instance, right- or left-wing extremism, also deserve attention. It is critical for the success of counterterrorism policies and social unity that the concerns are comprehensively analyzed regarding the ideologies that shape the views of the mass public and the political attitudes. Moreover, as terrorism is seen in many contexts other than the Western world, exploring this issue will contribute to a deeper global comprehension. Terrorism, being very much an issue based on the diversity of cultures, societies, and their political emotions, is not the same across the world. Years-long longitudinal studies which track individuals' perceptions over time will provide a wealth of information on the long-term effects of terrorism, to name some of these as a mention, radicalization, resilience, and social transformation. This brings up the fact that terrorists are always a constant factor in society with their potentially dangerous ideologies. Furthermore, there is the endless ethical and bias topic which is very sensitive and proposing lanes for future research on the same problem is well thought out.

\newpage

# Conclusion

Our study builds upon and extends the insightful work of Amélie Godefroidt [@Godefroidt2023], presenting new perspectives on the influence of terrorism on political attitudes. Through comprehensive analysis and extension of previous research, it highlights the significant ways in which terrorism leads shifts towards outgroup hostility, political conservatism, and affects public opinion across various contexts. This paper underscores the importance of considering the sociopolitical environment when evaluating the impact of terrorism and suggests a nuanced understanding of its effects is important for formulating policies. In moving forward, we suggest an inclusive approach to research that encompasses a broader spectrum of terrorist ideologies and embraces different methodologies to capture the influence of terrorism in our society. This approach not only contributes to the academic discussion on terrorism and democracy but also offers practical insights for policymakers to navigate the challenges posed by terrorism.

\newpage

# References